---
title: "Figures for 'Food processing and pottery use in the Early Bronze Age, Central Korean Peninsula'"
author: "Ben Marwick & Seunkgi Kwak"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This document reproduces the figures illustrating the geochemical data in Kwak and Marwick 'Food processing and pottery use in the Early Bronze Age, Central Korean Peninsula'

```{r}
library(kwakmarwickjas)
```

## Figures

The result chromatogram of the GC-MS analysis from one of our samples from the Sosa-Dong site (SOS049). Due to degradation, we usually observe medium- and long-chain saturated fatty acids. 5-α Cholestane was added as an internal standard (132 ng / microliter).   

```{r fig.height = 4, fig.width=8}
# Input data from csv file
SOS049 <- read.csv(file="data/GC_MS_exported_data_Figure4.csv", header=TRUE, sep=",")
make_retention_times_plot(SOS049)
```

The results of CSIA by GC-C-IRMS of the samples from the Sosa-Dong (ID: SO) and Kimpo-Yangchon (ID: KI) site using the reference from Craig et al., 2011 and Steele et al., 2010. 

```{r fig.height = 8, fig.width=8}
# read in data
CSIA_KM <- read.csv('data/CSIA_KM.csv', stringsAsFactors = FALSE)
CSIA_SS <- read.csv('data/CSIA_SS.csv', stringsAsFactors = FALSE)
# combine both into one dataframe
CSIA_KM_and_SS <- rbind(CSIA_KM, CSIA_SS)
# get site ID to control plotting colour and shape
CSIA_KM_and_SS$ID <- substr(CSIA_KM_and_SS$Sample.No., 1, 2)
# draw plot
make_C16_C18_scatter_plot(CSIA_KM_and_SS)
```
